---
title: "Hepatocytes and PKM2"
date: University of Bern, DBMR, Visceral and Transplantation Surgery lab
author: Tess Brodie
output: 
   html_document:
    toc_depth: 2
---

## Introduction

```{r setup, include = FALSE, dev='svg'}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(autodep = TRUE, cache = TRUE, cache.lazy = FALSE)
knitr::opts_chunk$set(dev = 'svg')
```



### Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(CATALYST)
library(dplyr)
library(flowCore)
library(ggplot2)
library(diffcyt)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(workflowr)
```

### Load data

```{r load-data}
sce <- readRDS(file.path("output", "sce.rds"))
```

## Liver subsets with PKM2 expression over time

```{r}
plotMultiHeatmap(sce, 
                 k = "mm", scale = "first",
                 hm1 = "type", hm2 = "PKM2",
                 row_anno = FALSE, col_clust = FALSE,
                 row_clust = FALSE, col_anno = FALSE,
                 bars = TRUE, perc = TRUE,
                 hm2_pal = hcl.colors(10, palette = "Reds"))

```

### Liver subsets with Unknown population taken out for better heatmap visualization

```{r}
sceU <- filterSCE(sce, k = "mm",
               cluster_id %in% c("B cells","BMDM", "cDCs",
                                 "Granulocytes","Hepatocytes", "KCs", "LSECs",
                                 "pDCs", "T cells"))
plotMultiHeatmap(sceU, 
                 k = "mm", scale = "first",
                 hm1 = "type", hm2 = "PKM2",
                 row_anno = FALSE, col_clust = FALSE,
                 row_clust = FALSE, col_anno = FALSE,
                 bars = TRUE, perc = TRUE,
                 hm2_pal = hcl.colors(10, palette = "Reds", rev = TRUE))

```
### Another color combo

```{r}
plotMultiHeatmap(sceU, 
                 k = "mm", scale = "first",
                 hm1 = "type", hm2 = "PKM2",
                 row_anno = FALSE, col_clust = FALSE,
                 row_clust = FALSE, col_anno = FALSE,
                 bars = TRUE, perc = TRUE)
```


## Hepatocytes and differential expression of PKM2
*need help scaling second tsne plot

```{r}
sceH <- filterSCE(sce, k = "mm",
               cluster_id %in% "Hepatocytes")

plot_grid(
  plotDR(sce, dr= "TSNE", color_by = "mm"),
  plotDR(sceH, dr= "TSNE", color_by = "mm"))
```

## tSNE of hepatocytes

```{r}
sceH <- runDR(sceH, dr = "TSNE", features = NULL)
plotDR(sceH, dr = "TSNE", color_by = "PKM2")

```
## PKM2 in Hepatocytes split by Condition

```{r}
plotDR(sceH, dr = "TSNE", color_by = "PKM2", facet_by = "condition")
```


## Proportion of hepatocytes per sample

```{r}
plotCounts(sceH, 
           prop = TRUE,
           group_by = "condition", 
           color_by = "patient_id")
```

## Number of hepatocytes per sample

```{r}
plotCounts(sceH, 
           group_by = "sample_id", 
           color_by = "condition")

```

## Hepatocyte subsets and PKM2 over time
We don't see one subset with all the PKM2

```{r}
sceH <- cluster(sceH, 
                xdim = 10, ydim = 10, maxK = 3, 
                verbose = FALSE)

plotMultiHeatmap(sceH, 
                 hm1 = "type", hm2 = "PKM2", 
                 k = "meta3", row_anno = FALSE,
                 col_dend = c(FALSE, TRUE))
```

### tSNE plot of hepatocyte subsets 

```{r}
plotDR(sceH, dr = "TSNE", color_by = "meta3")
```


### Cluster hepatocytes into many clusters to get a cluster with the PKM2 high cells in it
We still don't see one subset with all the PKM2

```{r}
sceH <- cluster(sceH, 
                xdim = 10, ydim = 10, maxK = 15, 
                verbose = FALSE)

plotMultiHeatmap(sceH, 
                 hm1 = "type", hm2 = "PKM2", 
                 k = "meta15", row_anno = FALSE,
                 col_dend = c(FALSE, TRUE))


```

### Plot clusters on tSNE

```{r}
plotDR(sceH, dr = "TSNE", color_by = "meta15")
```

